---
title: "CASA0005"
output: html_document
date: "2023-10-25"
---

# Hello!

## Welcome to this R notebook for CASA0005 module 🗺

This notebooks shows the steps to join the global gender inequality index to spatial data of the World, creating a new column of difference in inequality between 2010 and 2019, as per the instructions in the CASA0005 module for week 4 homework.

#### Hope you enjoy it! 😊

First, let's load some packages!

```{r message=FALSE, warning=FALSE}
library(dplyr) # we'll need that to join datasets
library(sf) # we'll need that to deal with spatial data
library(tmap)# we'll need that to make a map to check that what we did looks correct
library(countrycode)
```

Now, let's load some data!

```{r message=FALSE, warning=FALSE}
df <- read.csv("data/HDR21-22_Composite_indices.csv") # index data
sdf <- st_read("data/World_Countries_Generalized.geojson") # and spatial data
```

```{r}
df <- df[1:195, ] # only keep countries

df <- df |> 
  select(c(iso3, gii_2010,gii_2019,country)) |> 
  mutate(gii_diff = gii_2019 - gii_2010)

df$iso2 <- countrycode(df$iso3, origin = "iso3c", destination = "iso2c")

```

Joining dadasets on ISO2 codes

```{r}
sdf <- left_join(sdf, df, by = c("ISO" = "iso2"))
```

Mapping

```{r}
tm_shape(sdf) +
  tm_borders(lwd = 0.3) +
  tm_fill("gii_diff", palette = "Blues", style="jenks")
```

#### Hmmm, looks like there are quite a few missing values here. 😥

Let's see why that is
```{r}
not_matched <- df %>%
  anti_join(sdf, by = c("iso2" = "ISO")) |> # what countries didn't match correctly?
  filter(!is.na(gii_diff)) # and for which do we have the data?

not_matched # show the result
```

There are no countries which are not NA which haven't matched, so that's all good

## Choice! 🤩

Save the data if you need to.

```{r}
st_write(sdf, "genderInequalityIndex.gpkg")
```

Good luck with the Homework for week 5!

## and Happy Halloween 🎃
